---
title: "Take Home Exercise 1"
description: |
  In this exercise, we shall apply the skills we had learned in last two lessons to create data visualisation by using ggplot2.
author:
  - name: Reynard Lam 
    url: https://www.linkedin.com/in/reynard-lam-570644133/
    affiliation: School of Computing and Information Systems, SMU
    affiliation_url: https://scis.smu.edu.sg/
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# 1.0 Task 1

***Problem Statement: *** Create a pareto chart showing the distribution of returns by product sub-category. The data file to be used is the superstore-2021.xls file.


# 1.1 Installing and Launching R Packages

Before we get started, we will need to install the relevant R packages. 


```{r echo = TRUE}
packages = c('ggpubr', 'tidyverse', 'readxl', 'knitr')

for(p in packages){library
  if(!require(p, character.only = T)){
    install.packages(p)
  }

  library(p, character.only = T)
}
```

## 1.2 Importing Data

First, let us import the data into R by using *read_xls()*  of **readxl** package.  

```{r echo = TRUE}
orders <- read_xls("data/Superstore-2021.xls", sheet = "Orders")
returns <- read_xls("data/Superstore-2021.xls", sheet = "Returns")


```



# 1.3 Joining the two data frames


```{r echo = TRUE}
joined_tab <- left_join(returns, orders,
                        by = c('Order ID' = 'Order ID'))

freq_returned <- joined_tab %>% 
  count(`Sub-Category`) %>%
  rename(Returns = n)

freq_sorted <- freq_returned %>%
  arrange(desc(Returns))

freq_cum <- freq_sorted %>%
  mutate(cumfreq = cumsum(Returns))

freq_cum$`Sub-Category` <- factor(freq_cum$`Sub-Category`, levels = freq_cum$`Sub-Category`)
```


# 1.4 Creating the Pareto chart

```{r}
options(repr.plot.width=12, repr.plot.height=5)
ggplot(freq_cum, aes(x=`Sub-Category`, y=Returns)) + geom_bar(stat = "identity", aes(y=Returns), fill = 'lightgreen')+
  geom_point(aes(x=`Sub-Category`, y=cumfreq), color = 'grey20', pch=16, size=1) + 
  geom_path(aes(y=cumfreq, group=1), colour="grey20", size=0.9) +
  xlab('Sub-Category') +
  scale_y_continuous(
    name = 'Returns (absolute frequency)', breaks = seq(0,3500, by = 300), minor_breaks=seq(0,3500,100),
    sec.axis = sec_axis(~.*1/sum(freq_cum$Returns),name = 'Cumulative Frequency (%)', breaks = seq(0,1, by=0.1),
    labels= scales::percent_format()))+
    theme(panel.background = element_rect(fill = 'white', colour = 'grey60', size = 0.5, linetype = 'solid'),
          panel.grid.major = element_line(size = 0.8, linetype = 'solid', colour = 'grey85'),
          panel.grid.minor = element_line(size = 0.3, linetype = 'solid', colour = 'grey95'),
          text = element_text(size = 7.5)) +

    ggtitle("Pareto Chart of Returns with Cumulative Frequency by Sub-Category")
                        

   

```


# 2.0 Task 2

***Problem Statement: *** Create an age-sex pyramid showing the demographic structure of Singapore by age cohort and gender. For this task, the data titled *Singapore Residents by Planning Area / Subzone, Age Group, Sex and Type of Dwelling, June 2021* should be used.


# 2.1 Importing Data

Let us import the data into R by using *read.csv*  of **readr** package.  

```{r echo = TRUE}
popdata <- read.csv("data/respopagesextod2021.csv")


```

# 2.2 Cleaning of the data


```{r echo = TRUE}

popdata <- aggregate(formula = Pop ~ Sex + AG, data = popdata, FUN = sum)

## sort data by first by gender, then by age groups
popdata <- with(popdata, popdata[order(Sex,AG),])

## only use the three variables age, gender and population 
popdata <- popdata[,c("AG","Sex","Pop")]

## switch males to negative values so they appear on the left
popdata$Pop <- ifelse(popdata$Sex == "Males", -1*popdata$Pop, popdata$Pop)


## pyramid charts are two barcharts with axes flipped
ggplot(popdata, aes(x = AG, y = Pop, fill = Sex)) +
  geom_bar(data = subset(popdata, Sex == "Females"), stat = "identity") +
  geom_bar(data = subset(popdata, Sex == "Males"), stat = "identity") +
  scale_y_continuous(labels = paste0(as.character(c(seq(2, 0, -1), seq(1, 2, 1))), "m")) +
  coord_flip()

```

